---
title: "Final Analysis"
author: "Team 3 - Ryan and Anders"
date: "2/15/2021"
output: html_document
---

```{r setup}
choco <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/data/choco.csv") 
```

# 1. Rating Distribution
```{r}
hist(choco$Rating)
boxplot(choco$Rating)
summary(choco$Rating)
```
There are a few outliers at the 1.0 end of the spectrum.

# 2. Number of Ratings by Company
Since each row in the dataset represents one rating, we can just count the number
of rows for each company to get the number of ratings
```{r}
count_by_company <- sort(table(choco$Location), decreasing=TRUE)
barplot(count_by_company, ylab='Number of Ratings', las=2, ylim=c(0,1000))
mtext('Company Location', side=3, line=1)
```

We can see here just the 5 companies which have the most ratings:
```{r}
# Top 5
head(count_by_company)
```

# 3. Ratings by Cocoa Percent
We can visualize the mean ratings by cocoa percent with a box plot:
```{r}
boxplot(choco$Rating ~ choco$Cocoa_Percent, xlab='Cocoa Percent', ylab='Rating')
```

As seen above, the highest ratings generally come from cocoa percents in the range of 65-79%.

We can also show in a table the average cocoa percent for each rating interval:
```{r}
tb1 <- matrix(c(choco$Cocoa_Percent, choco$Rating), ncol=2)
aggregate(choco$Cocoa_Percent ~ choco$Rating, tb1, mean)
```

General as the rating increases, the average cocoa percent is close to 70%.

# 4. Bean Origin vs Rating
To get a general overview of what locations produce the cacao beans which taste
the best on average, we can box plot the rating property, categorizing by
bean origin:
```{r}

types <- reorder(choco$Bean_Origin, choco$Rating, FUN=mean)
boxplot(choco$Rating ~ types, las=2, ylab='Average Rating', xlab='')
mtext('Cacao Bean Origin', side=3, line=1)
```

However, if we want to know just where the best beans come from (i.e., beans
that produce chocolates which having a rating of 4.0), we can filter down
the dataset to only chocolates with the highest rating and show them in a table
or barplot:
```{r}
# Want bar chart of all 4.0 ratings by location
best <- choco[choco$Rating == 4.0,]
best_grouped <- sort(table(best$Bean_Origin))
best_grouped
barplot(best_grouped, las=2, ylab='Number of 4.0 Ratings', ylim=c(0,20))
mtext('Cacao Bean Origin', side=3, line=1)
```

So although Tobago produced on average the highest rated chocolates, we could guess
that this is because there are few data points i.e. chocolates tested from Tobago.

We can confirm this by showing how many chocolates were tested whose beans
originated from Tobago:
```{r}
table(choco$Bean_Origin)['Tobago']
```
Only 2 chocolates from Tobago were tested, which explains the extremely high average.

Venezuela on the other hand, as can be seen in the bar chart, produced beans
for 18 chocolates which were rated 4.0. We can see where Venezuela lies on average

```{r}
types <- reorder(choco$Bean_Origin, choco$Rating, FUN=function(x) -mean(x))
mean_ratings <- aggregate(choco$Rating ~ types, FUN=mean)
mean_ratings
```

We can see that Venezuela is rated 21st on average even though it has the most
4.0 ratings.

# 5. i. Influence of Rating Date on Rating
One analysis to perform could be if the date/year has an influence on rating. 
```{r}
boxplot(choco$Rating ~ choco$Date)
```

As you can see from the box plot, they haven't really improved the quality of chocolate as the years
have gone on. 

```{r}
tb1 <- matrix(c(choco$Date, choco$Rating), ncol=2)
aggregate(choco$Rating ~ choco$Date , tb1, mean)
```

As the years have gone on, they have gotten very consistent with the quality of chocolate, with a peak 
around 2017. 


# 5. ii. Influence of Taste Characteristics on Rating

First we can show some of the unique characteristics from ratings in the dataset:

```{r}
all_characteristics <- unique(c(choco$Characteristics1, choco$Characteristics2, 
                         choco$Characteristics3, choco$Characteristics4))
all_characteristics <- all_characteristics[!is.na(all_characteristics)]
str(all_characteristics)
```


A second analysis to perform would be how rating and Characteristic1 are related:
```{r}
x <- reorder(choco$Characteristics1, choco$Rating, FUN=mean)
boxplot(choco$Rating ~ x, las=2, xlab="")
```

This shows that chocolates with characteristics like "Tropical fruit" and "Sweet spice"
have the highest rating.

